Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-11-05 edubots_surveys.Rnw
Correcting students' survey

 The line of research of which this paper is a part has involved two phases of data collection.
 \begin{itemize}
\item The first phase aimed to collect feedback from students about their actual needs \cite{MoraChatbots2021:anon}. Hence, the objective was to collect information about students’ preferences of tools/applications for chatting and messaging, how they use them in educational context, who do they like to be with in the class messaging groups, and their expectations of the chatbots in terms of assisting them during their learning process. Therefore, two surveys for bachelor and master degree students at the University of Granada (Spain) were designed and answers from more than 250 students were collected. The key findings of the students surveys reveal that they prefer to use messaging applications, such as Telegram or WhatsApp as they are familiar to them. Students also expressed interest in chatbots that can assist them with organizing their course schedules, help them access their assignment grades and facilitate searching for course's resources. 
\item The results of the first phase were fed into the second phase of the study, which focused on teachers’ opinions about using chatbots and messaging applications in classes. This is what is going to be presented in this paper.
\end{itemize}

Initially, a single survey was developed for this purpose. Responses from 300 teachers were collected and analysed. An initial analysis of these responses led to the creation of a second survey, which collected responses from 200 different higher education teachers. The questions on this survey focus on discussing the implications and the impact of the development of future technology-enhanced tools and educational institutions’ policies by finding answers to the following research questions:

\begin{itemize}
\item RQ1 - Are teachers already using messaging apps in their classes? 
\item RQ2 - Which kind of chatbots would teachers find useful in their classes? 
\item RQ3 - Which kind of interaction do teachers prefer with their students?
\item RQ4 - What kind of interaction media features do teachers value the most?
\end{itemize}

@@

% The Use of Chatbots in Higher Education
The use of MIM, although possibly valuable by itself, can be enhanced via the use of chatbots, which, being conversational agents, usually dwell in systems where synchronous conversations take place. 
The use of conversational agents (chatbots) in higher education is
still at its infancy \cite{yang2019opportunities}. Nevertheless,
recent studies examining their positive impact on students' academic
performance \cite{perez2020rediscovering} and engagement
\cite{differ.chat,abbas2021onlinechat} have led to a growing interest
in using this technology in the (possibly virtual) classroom. Indeed, the use of chatbots in
collecting course feedback from students in higher education improved
students' response quality and boosted their enjoyment levels
\cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, using
either tools such as mobile devices or teaching strategies based on
gamification \cite{Yildirim2017} can improve student motivation. In
this sense, the authors in \cite{Pimmer2019} carried out a series of experiments with
technological tools whose results showed positive perception and
acceptance of the use of technology in teaching and learning.

Several higher education chatbots' evaluation studies have been 
undertaken. For instance, a recent evaluation review study presented by Smutny and Schreiberova 
\cite{Smutny2020} examined 47 educational chatbots implemented in
Facebook Messenger with the focus to identify characteristics and
quality metrics such as language, subject matter and platform. 
Whereas study \cite{perez2020rediscovering} aimed to categorise educational chatbots according to their
purpose into service-oriented and teaching-oriented. The first category includes those that
provide service support such as the chatbot Ask Holly \cite{meetholly}
and Dina \cite{santoso2018dinus}; both chatbots respond to students'
questions about enrolment and registration. Ask L. U.
 \cite{LancasterUniversity_2019} answers students' frequently
asked questions about timetables, grades, tutors and societies. LISA
\cite{dibitonto2018chatbot} and Differ \cite{differ.chat} facilitate
breaking the ice between new students by introducing them to each
other. Ranoliya et al. \cite{ranoliya2017chatbot} proposed a generic chatbot for
university students that is able to answer to their frequently asked
questions. 
For some time, the University of Granada had ELVIRA as a
chatbot in its main web page \cite{MOREO20129}. Besides being able to
answer pre-established (frequent) questions, it did so from the website of the
University of Granada using an inset character, voiced by a real
person, who lip-synced the answers. The emphasis in the case of Elvira,
however, was on the authenticity of the synchronization more than in
making the update of questions and answers simpler, or more
interactive, using actual messaging platforms. 
In most cases, using the inset search engine was able to give you more up-to-date and
accurate answers than the ones provided by ELVIRA, which had to be updated by hand (and not too often).

@@

Regarding the factors for the adoption of chatbots in higher
education, many studies have focused on the evaluation of technology
acceptance and usability \cite{Roblyer2010,Pimmer2019}. However,
higher education is a special domain where, according to Hobert \cite{hobert2019you}, 
specific pedagogical factors such as learning success or increased motivation are more important.  
Therefore, to develop efficient chatbots for higher education, all stakeholders'
(i.e. educators, students, institutions, etc) expectations needs to be
carefully collected and taken into consideration
\cite{sjostrom2018designing,tsivitanidouusers}.

@@

%Why messaging apps are relevant here --> connect to the development of chatbots
Group messaging apps have become a powerful tool for keeping touch with students. Also, they are the baseline platform for the development of chatbots to help with the learning process. After an initial analysis to the answers to the first survey, we realized that there were some prior issues mainly related to the adoption of a technology, chatbots, that generally piggybacks on another, messaging applications. In a technology adoption model, perceived ease of use and perceived usefulness are essential. But it is going to be very difficult for teachers to find {\em chatbots} easy to use if they, previously, do not consider, or simply use, the messaging tools to which they are attached.

This is why for answering RQ1, teachers were queried about whether they use messaging apps in their classes, specifically, Telegram, WhatsApp, Slack (an application used mainly in IT departments and software development), or any other messaging app, and whether they use messaging apps provided by their academic institution (see Table \ref{tab:use_messapps}). In this case, it is maybe in order a certain clarification. While we had in mind, as indicated by the possible answers, instant messaging applications when making this question, we included these institution-provided applications mainly for completitude. But it should be noted that, in most cases, they are not {\em instant} messaging apps in the way Telegram or Facebook Messenger are; in most cases they are messaging facilities provided by learning management systems such as (usually) Moodle. So the answer to this question must be understood in two different ways: their preference for the {\em source} of the technology (university, any company) as well as the {\em type} of technology (instant, app based messaging systems vs. web-based systems integrated in another learning management system).

>>>>

 The line of research of which this paper is a part has involved two phases of data collection.
 \begin{itemize}
\item The first phase aimed to collect feedback from students about their actual needs \cite{MoraChatbots2021:anon}. Hence, the objective was to collect information about students’ preferences of tools/applications for chatting and messaging, how they use them in educational context, who do they like to be with in the class messaging groups, and their expectations of the chatbots in terms of assisting them during their learning process. Therefore, two surveys for bachelor and master degree students at the University of Granada (Spain) were designed and answers from more than 250 students were collected. The key findings of the students' surveys reveal that they prefer to use messaging applications, such as Telegram or WhatsApp as they are familiar to them. Students also expressed interest in chatbots that can assist them with organizing their course schedules, help them access their assignment grades and facilitate searching for course's resources. 
\item The results of the first phase were fed into the second phase of the study, which focused on teachers’ opinions about using chatbots and messaging applications in classes. This is what is going to be presented in this paper.
\end{itemize}

Initially, a single survey was developed for this purpose and responses from 300 higher education teachers were collected and analysed. 
The responses' analysis entailed the creation of a second survey to which 200 teachers' responses were collected. The questions of the second 
survey focussed on discussing teachers' needs, opinions and preferences of the development of future technology-enhanced tools 
and its potential impact on educational institutions’ policies. Therefore, this paper investigated the following research questions:

\begin{itemize}
\item RQ1 - Are teachers already using messaging apps in their classes? 
\item RQ2 - Which chatbots' features would teachers find useful in their classes? 
\item RQ3 - Which kind of interaction do teachers prefer with their students?
\item RQ4 - What kind of interaction media features do teachers value the most?
\end{itemize}

@@

% The Use of Chatbots in Higher Education
The use of MIM, although possibly valuable by itself, can be enhanced via the use of chatbots, which, being conversational agents, usually dwell
in systems where synchronous conversations take place. 
The use of conversational agents (chatbots) in higher education is
still at its infancy \cite{yang2019opportunities}. Nevertheless,
recent studies examining their positive impact on students' academic
performance \cite{perez2020rediscovering} and engagement
\cite{differ.chat,abbas2021onlinechat} have led to a growing interest
in using this technology in the (possibly virtual) classroom. Indeed, the use of chatbots in
collecting course feedback from students in higher education improved
students' response quality and boosted their enjoyment levels
\cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, using
either tools such as mobile devices or teaching strategies based on
gamification \cite{Yildirim2017} can improve student motivation. In
this sense, the authors in \cite{Pimmer2019} adopted a quasi-experimental, survey-based approach to
report the positive impact of using instant messaging tools in boosting students' knowledge and mitigating their feelings of isolation.

Several higher education chatbots' evaluation studies have been 
undertaken. For instance, a recent evaluation review study presented by Smutny and Schreiberova 
\cite{Smutny2020} examined 47 educational chatbots implemented in
Facebook Messenger with the focus to identify characteristics and
quality metrics such as language, subject matter and platform. 
Whereas the study undertaken by \cite{perez2020rediscovering} aimed to categorise educational chatbots according to their
purpose into service-oriented and teaching-oriented. The first category includes those that
provide service support such as the chatbot Ask Holly \cite{meetholly}
and Dina \cite{santoso2018dinus}; both chatbots respond to students'
questions about enrolment and registration. Ask L. U.
 \cite{LancasterUniversity_2019} answers students' frequently
asked questions about timetables, grades, tutors and societies. LISA
\cite{dibitonto2018chatbot} and Differ \cite{differ.chat} facilitate
breaking the ice between new students by introducing them to each
other. Ranoliya et al. \cite{ranoliya2017chatbot} proposed a generic chatbot for
university students that is able to answer to their frequently asked
questions. 
For some time, the University of Granada had ELVIRA as a
chatbot in its main web page \cite{MOREO20129}. Besides being able to
answer pre-established (frequent) questions, it did so from the website of the
University of Granada using an inset character, voiced by a real
person, who lip-synced the answers. The emphasis in the case of Elvira,
however, was on the authenticity of the synchronization more than in
making the update of questions and answers simpler, or more
interactive, using actual messaging platforms. 
In most cases, using the inset search engine was able to give you more up-to-date and
accurate answers than the ones provided by ELVIRA, which had to be updated by hand (and not too often).

@@

Regarding the factors for the adoption of chatbots in higher
education, many studies have focused on the evaluation of technology
acceptance and usability \cite{Roblyer2010,Pimmer2019}. However,
higher education is a special domain where, according to Hobert \cite{hobert2019you}, 
specific pedagogical factors such as learning success or increased motivation are more important.  
Therefore, to develop efficient chatbots for higher education, all stakeholders'
(i.e. educators, students, institutions, etc) needs should be
carefully collected and taken into consideration
\cite{sjostrom2018designing,tsivitanidouusers}.

@@

%Why messaging apps are relevant here --> connect to the development of chatbots
Group messaging apps have become a powerful tool for keeping touch with students. Also, many of them have built-in chatbots 
to enhance the learning process. After an initial analysis to the answers to the first survey, 
we realized that there were some prior issues mainly related to the adoption of a technology, chatbots, that generally 
piggybacks on another, messaging applications. In a technology adoption model, perceived ease of use and perceived usefulness are essential. But it is going to be very difficult for teachers to find {\em chatbots} easy to use if they, previously, do not consider, or simply use, the messaging tools to which they are attached.

This is why for answering RQ1, teachers were queried about whether they use messaging apps in their classes, specifically, Telegram, WhatsApp, Slack (an application used mainly in IT departments and software development), or any other messaging app, and whether they use messaging apps provided by their academic institution (see Table \ref{tab:use_messapps}). In this case, it is maybe in order a certain clarification. While we had in mind, as indicated by the possible answers, instant messaging applications when making this question, we included these institution-provided applications mainly for completion. But it should be noted that, in most cases, they are not {\em instant} messaging apps in the way Telegram or Facebook Messenger are; in most cases they are messaging facilities provided by learning management systems such as (usually) Moodle. So the answer to this question must be understood in two different ways: their preference for the {\em source} of the technology (university, any company) as well as the {\em type} of technology (instant, app based messaging systems vs. web-based systems integrated in another learning management system).
